Active Learning for Level Set Estimation

### 2013-03-22 msthpres.tex
Revert to sans serif; fixes in algorithm
More fixes all around

\documentclass[10pt,mathserif,serif]{beamer}
%\documentclass[10pt,mathserif,serif,handout]{beamer}

@@

\begin{frame}
\begin{center}
\uncover<1->{Swimmers of Lake Zurich, beware!}
\vspace{0.2in}
\begin{columns}[c]
\column{0.4\textwidth}
\uncover<1->{
\sig{\includegraphics[width=1.7in]{figures/pr.jpeg}}{www.limnobotics.ch}\\[1em]
\sig{\includegraphics[width=1.7in]{figures/bloom.jpg}}{Flickr/Dr. Jennifer L. Graham/U.S. Geological Survey}
}
\column{0.6\textwidth}
\minipage[c][0.80\textheight][s]{\columnwidth}
\uncover<1->{
\small
``The warming waters of one of central Europe's most popular holiday
destinations, Switzerland's Lake Zurich, have created an ideal environment
for a population explosion of algae including \emph{Planktothrix rubescens},
[...]''
\qauth{Scientific American}
}
\vfill
\uncover<2->{
\small
``\emph{Planktotrhix rubescens} can account for half of the total
phytoplankton biomass in Lake Zurich in summer [...]''
\vspace{1em}

``\emph{Planktotrhix rubescens} are among the most important
producers of hepatotoxic microcystins in freshwaters [...]''
\qauth{Silke Van den Wyngaert \emph{et al.}, ASLO, 2011}
}
\vfill
\uncover<3->{
\small
``Microcystins [...] are cyanotoxins and can be very toxic for plants and
animals including humans.
Their hepatotoxicity may cause serious damage to the liver.''
\qauth{Wikipedia}
}
\endminipage
\end{columns}
\end{center}
\end{frame}

@@

\begin{frame}
\begin{center}
\uncover<1->{\large Why GPs are nice}
\end{center}
\begin{itemize}
\item<2-> \textbf{Nonparametric}:\\
      Impose ``smoothness'' assumptions via kernel $k(x, x')$.
\vspace{1em}
\item<3-> \textbf{Bayesian}, yet \textbf{efficient}:
      \begin{align*}
        \mu_t(\*x) &= \*k_t(\*x)^T\left(\*K_t + \sigma^2 \*I\right)^{-1}\*y_t\\
        k_t(\*x, \*x') &= k(\*x, \*x') - \*k_t(\*x)^T\left(\*K_t + \sigma^2 \*I\right)^{-1}\*k_t(\*x)\notag\\
        \sigma_t^2(\*x) &= k_t(\*x, \*x)
      \end{align*}
      Suitable for step-by-step updates.
\vspace{1em}
\item<4-> More importantly: they provide \textbf{variance estimates}!\\
          Construct confidence intervals: $Q_t(\*x) = \left[\mu_{t-1}(\*x) \pm \beta_t^{1/2} \sigma_{t-1}(\*x)\right]$
\end{itemize}
\end{frame}

@@

\begin{frame}
\begin{columns}[t]
\column{0.5\textwidth}
{\footnotesize
\begin{algorithmic}[1]
  \REQUIRE sample set $D$, GP prior ($\mu_0 = 0$, $k$, $\sigma_0$),\\
           \hspace{1.35em}threshold value $h$, accuracy parameter $\epsilon$
  \ENSURE predicted sets $\hat{H}$, $\hat{L}$
  \uncover<2->{
  \CSTATE{c1}$H_0 \gets \varnothing$,\enskip $L_0 \gets \varnothing$,\enskip $U_0 \gets D$ \label{lin:init1}
  %\LET{$H_0$}{$\varnothing$} \label{lin:init1}
  %\LET{$L_0$}{$\varnothing$}
  %\LET{$U_0$}{$D$}
  \CLET{c1}{$C_0(\*x)$}{$\mathbb{R}$, for all $\*x \in D$} \label{lin:init2}
  \CLET{c1}{$t$}{1}
  }
  \uncover<3->{
  \CWHILE{c6}{$U_{t-1} \neq \varnothing$}
    \CSTATE{c6}$H_t \gets H_{t-1}$,\enskip $L_t \gets L_{t-1}$,\enskip $U_t \gets U_{t-1}$
    }
    %\LET{$H_t$}{$H_{t-1}$}
    %\LET{$L_t$}{$L_{t-1}$}
    %\LET{$U_t$}{$U_{t-1}$}
    \uncover<4->{
    \CFORALL{c2}{$\*x \in U_{t-1}$}
      \CLET{c2}{$C_{t}(\*x)$}{$C_{t-1}(\*x) \cap Q_t(\*x)$} \label{lin:upd}
      \CIF{c2}{$\min(C_t(\*x)) + \epsilon > h$} \label{lin:class1}
        \CLET{c2}{$U_t$}{$U_t \setminus \{\*x\}$}
        \CLET{c2}{$H_t$}{$H_t \cup \{\*x\}$} 
      \CELSIF{c2}{$\max(C_t(\*x)) - \epsilon \leq h$} \label{lin:classr2}
        \CLET{c2}{$U_t$}{$U_t \setminus \{\*x\}$}
        \CLET{c2}{$L_t$}{$L_t \cup \{\*x\}$}
      \CENDIF{c2} \label{lin:class2}
    \CENDFOR{c2}
    }
    \uncover<5->{
    \CLET{c3}{$\*x_t$}{$\argmax_{\*x \in U_t}(a_t(\*x))$} \label{lin:sel1}
    \CLET{c3}{$y_t$}{$f(\*x_t) + n_t$} \label{lin:sel2}
    }
    \uncover<6->{
    \CSTATE{c4}Compute $\mu_t(\*x)$ and $\sigma_t(\*x)$, for all $\*x \in U_t$ \label{lin:inf}
    }
    \uncover<3->{
    \CLET{c6}{$t$}{$t + 1$}
  \CENDWHILE{c6}
  }
  \uncover<2->{
  \CLET{c5}{$\hat{H}$}{$H_{t-1}$} \label{lin:ret1}
  \CLET{c5}{$\hat{L}$}{$L_{t-1}$} \label{lin:ret2}
  }
\end{algorithmic}
}
\column{0.5\textwidth}
\begin{center}
The Level Set Estimation (\acl) algorithm
\end{center}
\only<3-6>{\vspace{34pt}$\leftarrow$ {\small loop until all points have been classified\\}\vspace{40pt}}
\only<4-6>{$\leftarrow$ {\small classify\\}\vspace{52pt}}
\only<5-6>{$\leftarrow$ {\small select max. ambiguity point\\}\vspace{3pt}}
\only<6>{$\leftarrow$ {\small update GP estimate}}
\only<7->{
\vspace{3pt}
\begin{itemize}
\only<7->{
\item Monotonicity of
\begin{enumerate}
\item confidence intervals
\item classification
\end{enumerate}
}
\vspace{0em}
\only<8->{
\item Relaxed classification rules by an accuracy parameter $\epsilon$\\
      (trades off sampling cost for accuracy)
}
\end{itemize}
}
\begin{center}
\includegraphics<8->[width=2in]{figures/voned_cl_41_eps}
\end{center}
\end{columns}
\end{frame}

@@

 \uncover<11->{
 \node[st,left=2em of amb] (t0) {Monotonicity}
   edge[pil] (amb);
 }

@@

 \uncover<15->{
 \node[st,right=2em of loss] (rules) {Class. rules}
   edge[pil] (loss);
 }

@@

 \uncover<16->{ 
 \node[st,right=2em of est] (delta) {$\delta$}
   edge[pil] (est);
 }

@@

\begin{frame}
\begin{itemize}
\item<1-> Up to this point we have assumed a fixed cost per sample
\item<2-> What about the traveling distance between measurements?
\item<3-> Plan ahead:
\begin{itemize}
\item Select a \emph{batch} of sampling locations
\item Connect them using a Euclidean TSP path
\item Traverse the path and collect measurements
\end{itemize}
\end{itemize}

@@

\begin{frame}
\begin{center}
\large Summary
\end{center}
\vspace{-1em}
\begin{itemize}
\item<1-> \acl algorithm:
\begin{itemize}
\item<2-> \begin{tabular}{m{5.5cm} >{\centering}m{4cm}}Theoretical guarantees & \includegraphics[width=1.3in]{figures/thm.png} \end{tabular}
\item<3-> \begin{tabular}{m{5.5cm} >{\centering}m{4cm}}Competitive with the state-of-the-art in practice (sometimes considerably superior) & \hspace{-0.5em}\includegraphics[width=1.3in]{figures/ev_bgape_seq_long} \end{tabular}
\end{itemize}
\item<4-> Two useful extensions:
\begin{itemize}
\item<5-> \begin{tabular}{m{5.5cm} >{\centering}m{4cm}}Implicit threshold level & \includegraphics[width=1.3in]{figures/limno_bgape_imp_class_140} \end{tabular}
\item<6-> \begin{tabular}{m{5.5cm} >{\centering}m{4cm}}Batch sampling $\rightarrow$ path planning & \includegraphics[width=1.3in]{figures/limno_bgape_pp_30_2} \end{tabular}
\end{itemize}
\item<7-> Look out for algae when swimming in Lake Zurich! =)
\end{itemize}
\end{frame}

>>>>

\documentclass[10pt,mathserif]{beamer}
%\documentclass[10pt,mathserif,handout]{beamer}

@@

\begin{frame}
\begin{center}
\uncover<1->{Swimmers of Lake Zurich, beware!}
\vspace{0.2in}
\begin{columns}[c]
\column{0.4\textwidth}
\uncover<2->{
\sig{\includegraphics[width=1.7in]{figures/pr.jpeg}}{www.limnobotics.ch}\\[1em]
\sig{\includegraphics[width=1.7in]{figures/bloom.jpg}}{Flickr/Dr. Jennifer L. Graham/U.S. Geological Survey}
}
\column{0.6\textwidth}
\minipage[c][0.80\textheight][s]{\columnwidth}
\uncover<1->{
\small
``The warming waters of one of central Europe's most popular holiday
destinations, Switzerland's Lake Zurich, have created an ideal environment
for a population explosion of algae including \emph{Planktothrix rubescens},
[\ldots]''
\qauth{Scientific American}
}
\vfill
\uncover<3->{
\small
``\emph{Planktotrhix rubescens} can account for half of the total
phytoplankton biomass in Lake Zurich in summer [\ldots]''
\vspace{1em}

``\emph{Planktotrhix rubescens} are among the most important
producers of hepatotoxic microcystins in freshwaters [\ldots]''
\qauth{Silke Van den Wyngaert \emph{et al.}, ASLO, 2011}
}
\vfill
\uncover<4->{
\small
``Microcystins [\ldots] are cyanotoxins and can be very toxic for plants and
animals including humans.
Their hepatotoxicity may cause serious damage to the liver.''
\qauth{Wikipedia}
}
\endminipage
\end{columns}
\end{center}
\end{frame}

@@

\begin{frame}
\begin{center}
\uncover<1->{\large Why GPs are nice}
\end{center}
\begin{itemize}
\item<2-> \textbf{Nonparametric}:\\
      Impose ``smoothness'' assumptions via kernel $k(x, x')$.
\vspace{1em}
\item<3-> \textbf{Bayesian}, yet \textbf{efficient}:
      \begin{align*}
        \mu_t(\*x) &= \*k_t(\*x)^T\left(\*K_t + \sigma^2 \*I\right)^{-1}\*y_t\\
        k_t(\*x, \*x') &= k(\*x, \*x') - \*k_t(\*x)^T\left(\*K_t + \sigma^2 \*I\right)^{-1}\*k_t(\*x)\notag\\
        \sigma_t^2(\*x) &= k_t(\*x, \*x)
      \end{align*}
      Suitable for step-by-step updates.
\vspace{1em}
\item<4-> They provide \textbf{variance estimates}!\\
          Construct confidence intervals: $Q_t(\*x) = \left[\mu_{t-1}(\*x) \pm \beta_t^{1/2} \sigma_{t-1}(\*x)\right]$
\end{itemize}
\end{frame}

@@

\begin{frame}
\begin{columns}[t]
\column{0.45\textwidth}
{\footnotesize
\begin{algorithmic}
  \REQUIRE sample set $D$, GP prior ($\mu_0$, $k$, $\sigma_0$),\\
           \hspace{2.2em}thr. value $h$, accuracy parameter $\epsilon$
  \ENSURE predicted sets $\hat{H}$, $\hat{L}$
  \uncover<2->{
  \CSTATE{c1}$H_0 \gets \varnothing$,\enskip $L_0 \gets \varnothing$,\enskip $U_0 \gets D$ \label{lin:init1}
  %\LET{$H_0$}{$\varnothing$} \label{lin:init1}
  %\LET{$L_0$}{$\varnothing$}
  %\LET{$U_0$}{$D$}
  \CLET{c1}{$C_0(\*x)$}{$\mathbb{R}$, for all $\*x \in D$} \label{lin:init2}
  \CLET{c1}{$t$}{1}
  }
  \uncover<3->{
  \CWHILE{c6}{$U_{t-1} \neq \varnothing$}
    \CSTATE{c6}$H_t \gets H_{t-1}$,\enskip $L_t \gets L_{t-1}$,\enskip $U_t \gets U_{t-1}$
    }
    %\LET{$H_t$}{$H_{t-1}$}
    %\LET{$L_t$}{$L_{t-1}$}
    %\LET{$U_t$}{$U_{t-1}$}
    \uncover<4->{
    \CFORALL{c2}{$\*x \in U_{t-1}$}
      \CLET{c2}{$C_{t}(\*x)$}{$C_{t-1}(\*x) \cap Q_t(\*x)$} \label{lin:upd}
      \CIF{c2}{$\min(C_t(\*x)) + \epsilon > h$} \label{lin:class1}
        \CLET{c2}{$U_t$}{$U_t \setminus \{\*x\}$}
        \CLET{c2}{$H_t$}{$H_t \cup \{\*x\}$} 
      \CELSIF{c2}{$\max(C_t(\*x)) - \epsilon \leq h$} \label{lin:classr2}
        \CLET{c2}{$U_t$}{$U_t \setminus \{\*x\}$}
        \CLET{c2}{$L_t$}{$L_t \cup \{\*x\}$}
      \CENDIF{c2} \label{lin:class2}
    \CENDFOR{c2}
    }
    \uncover<5->{
    \CLET{c3}{$\*x_t$}{$\argmax \{a_t(\*x)\mid \*x \in U_t\}$} \label{lin:sel1}
    \CLET{c3}{$y_t$}{$f(\*x_t) + n_t$} \label{lin:sel2}
    }
    \uncover<6->{
    \CSTATE{c4}Compute $\mu_t(\*x)$ and $\sigma_t(\*x)$, $\forall\*x \in U_t$ \label{lin:inf}
    }
    \uncover<3->{
    \CLET{c6}{$t$}{$t + 1$}
  \CENDWHILE{c6}
  }
  \uncover<2->{
  \CLET{c5}{$\hat{H}$}{$H_{t-1}$} \label{lin:ret1}
  \CLET{c5}{$\hat{L}$}{$L_{t-1}$} \label{lin:ret2}
  }
\end{algorithmic}
}
\column{0.55\textwidth}
\begin{center}
The Level Set Estimation (\acl) algorithm
\end{center}
\only<3-6>{\vspace{34pt}$\leftarrow$ {\small loop until all points have been classified\\}\vspace{40pt}}
\only<4-6>{$\leftarrow$ {\small classify\\}\vspace{52pt}}
\only<5-6>{$\leftarrow$ {\small select max. ambiguity point\\}\vspace{3pt}}
\only<6>{$\leftarrow$ {\small update GP estimate}}
\only<7->{
\vspace{3pt}
\begin{itemize}
\only<7->{
\item Monotonicity of
\begin{enumerate}
\item confidence intervals
\item classification
\end{enumerate}
}
\vspace{0em}
\only<8->{
\item Relaxed classification rules by an accuracy parameter $\epsilon$\\
      (trades off sampling cost for accuracy)
}
\end{itemize}
}
\begin{center}
\includegraphics<8->[width=2in]{figures/voned_cl_41_eps}
\end{center}
\end{columns}
\end{frame}

@@

 \uncover<11->{
 \node[st,left=1.5em of amb] (t0) {Monotonicity}
   edge[pil] (amb);
 }

@@

 \uncover<15->{
 \node[st,right=1.5em of loss] (rules) {Class. rules}
   edge[pil] (loss);
 }

@@

 \uncover<16->{ 
 \node[st,right=1.5em of est] (delta) {$\delta$}
   edge[pil] (est);
 }

@@

\begin{frame}
\begin{itemize}
\item<1-> Up to this point we have assumed a fixed cost per sample
\item<2-> What about the traveling distance between measurements?
\item<3-> Plan ahead:
\begin{itemize}
\item Select a \emph{batch} of sampling locations
\item Connect them using a Euclidean TSP path
\item Traverse path and collect measurements
\end{itemize}
\end{itemize}

@@

\begin{frame}
\begin{center}
\large Summary
\end{center}
\vspace{-1em}
\begin{itemize}
\item<1-> \acl algorithm:
\begin{itemize}
\item<2-> \begin{tabular}{m{5.7cm} >{\centering}m{4cm}}Theoretical guarantees & \includegraphics[width=1.3in]{figures/thm.png} \end{tabular}
\item<3-> \begin{tabular}{m{5.7cm} >{\centering}m{4cm}}Competitive with the state-of-the-art in practice (sometimes considerably superior) & \hspace{-0.5em}\includegraphics[width=1.3in]{figures/ev_bgape_seq_long} \end{tabular}
\end{itemize}
\item<4-> Two useful extensions:
\begin{itemize}
\item<5-> \begin{tabular}{m{5.7cm} >{\centering}m{4cm}}Implicit threshold level & \includegraphics[width=1.3in]{figures/limno_bgape_imp_class_140} \end{tabular}
\item<6-> \begin{tabular}{m{5.7cm} >{\centering}m{4cm}}Batch sampling $\rightarrow$ path planning & \includegraphics[width=1.3in]{figures/limno_bgape_pp_30_2} \end{tabular}
\end{itemize}
\item<7-> Look out for algae when swimming in Lake Zurich! =)
\end{itemize}
\end{frame}
